---
title: "Code to process all indicators and create final data frame for sampled California residential areas"
author: "Jackson Phippen"
date: "`r Sys.Date()`"
output: 
  html_document: 
    self_contained: false
---

## Process Indicators

Currently (as of June 1st, 2026), there are three subsets of the same sample:

1.  **All 5,000** sampled points via population weighting
2.  Thinned sampled points to be at least **1km** from all prisons and neighboring points
3.  Thinned sampled points to be at least **5km** from all prisons and neighboring points

To avoid weird computing, I will compute indicators for all 5,000 points and correctly subset points for future analysis.

There are 11 indicators in this code base. The majority of which are calculated in R, and have their own functions in the `R/` folder. Two indicators (canopy cover and heat exposure) are calculated via Google Earth Engine and have stand alone Python scripts in the 'python/' folder.

Component scores will not be calculated for these data. Use the function `load_ca_samples` in the `analysis/` folder to return the thinned subsets. 

First set up the environment with all required libraries via "setup.R" file, and load in data of interest.

```{r}
# setup environment
source("setup.R")

# read in sampled points and california prisons
ca_residences <- read_sf("data/ca_sample/all_ca_residences.gpkg")
thin_1km <- read_sf("data/ca_sample/1km_ca_residences.gpkg")
thin_5km <- read_sf("data/ca_sample/5km_ca_residences.gpkg")

ca_prisons <- read_sf("data/phase2/processed/prisons/study_prisons.gpkg") %>% 
  filter(STATE == "CA")

#manipulate order of ca_residences and drop cell and probabilities
ca_residences <- ca_residences %>% 
  select(object_id, geom)
```

#### Flood Risk

```{r}
# flood risk (takes 1-2 days to run) (Last run time was 1.72 hours, not sure how/why)
flood_risk <- calc_flood_risk(
  sf_obj = ca_residences,
  dist = 1000,
  id_column = "object_id",
  save = TRUE,
  out_path = "outputs/comparison/"
)
```

#### Wildfire Risk

```{r}
# wildfire risk
wildfire_risk <- calc_wildfire_risk(
  sf_obj = ca_residences,
  file = "data/phase2/raw/wildfire_risk/Data/whp2023_GeoTIF/",
  dist = 1000,
  conus_only = TRUE,
  id_column = "object_id",
  save = TRUE,
  out_path = "outputs/comparison/"
)
```

#### Ozone

```{r}
# ozone
ozone <- calc_ozone(
  sf_obj = ca_residences, 
  years    = c(2019, 2021, 2022),
  folder   = "data/phase2/raw/ozone",
  id_column = "object_id",
  save     = TRUE,
  out_path = "outputs/comparison/"
)
```

#### PM 2.5

```{r}
# pm2.5 (may change dataset, check for more recent years)
pm25 <-
  calc_pm25(
    sf_obj = ca_residences,
    folder = "data/phase2/processed/PM2.5",
    dist = 1000,
    years = c(2021, 2023),
    id_column = "object_id",
    save = TRUE,
    out_path = "outputs/comparison/"
  )
```

#### Pesticides

```{r}
pesticides <- calc_pesticides(
  ca_residences,
  folder = "data/phase2/raw/pesticide/PEST-CHEMGRIDS_v2/NC/",
  dist = 0, # resolution is 5km
  id_column = "object_id",
  save = TRUE,
  out_path = "outputs/comparison/")
```

#### Traffic Proximity

```{r}
#takes 1.5 days to run
traffic_prox <- calc_traffic_proximity(
  sf_obj = ca_residences,
  file = "data/phase2/processed/traffic/aadt_2023.RData",
  dist = 500,
  id_column = "object_id",
  save = TRUE,
  out_path = "outputs/comparison/"
)
```

#### Risk Management Plan (RMP) Facility Proximity

```{r}
#calculate Risk Management Plan (RMP) facility proximity
rmp_prox <- calc_rmp_proximity(
  sf_obj = ca_residences,
  file = "data/phase2/raw/RMP/EPA RMP Spreadsheets — The Data Liberation Project - Facilities.csv",
  dist = 5000,
  id_column = "object_id",
  save = TRUE,
  out_path = "outputs/comparison/"
)
```

#### NPL/Superfund Facility Proximity

```{r}
# calculate NPL facility proximity
npl_prox <- calc_npl_proximity(
  sf_obj = ca_residences,
  file = "data/phase2/processed/NPL/NPL_arc_geocode.shp",
  dist = 5000,
  id_column = "object_id",
  save = TRUE,
  out_path = "outputs/comparison/"
)
```

#### Hazardous Waste Facility Proximity

```{r}
# calculate Haz waste facility proximity
haz_prox <- calc_haz_waste_proximity(
  sf_obj = ca_residences,
  file = "data/phase2/processed/hazardous_waste/TSD_LQGs.csv",
  dist = 5000,
  id_column = "object_id",
  save = TRUE,
  out_path  = "outputs/comparison/"
)
```

### Final Data Frame

Because many indicators take a long time to run, let's read in each of the indicators. (In other words, it may be unrealistic to generate each indicator in a single session.) Be sure to read in indicators from GEE as well--note these updates are held in `data/ca_sample/`.

```{r}
# read in indicators 
# climate scores: 
flood_risk <- read_csv("outputs/comparison/flood_risk_2026-06-01.csv")

wildfire_risk <- read_csv("outputs/comparison/wildfire_risk_2026-06-01.csv")

heat_exp <- read_csv("data/ca_sample/ca_comparison_lst_daily_MODIS_2026-06-17.csv") %>% 
  group_by(object_id) %>%
  summarise(lst_avg = median(LST_mean, na.rm = TRUE))

canopy_cover <- read_csv("data/ca_sample/ca_sample_canopy_CONUS.csv")%>%
  rename(percent_tree_cover = NLCD_Percent_Tree_Canopy_Cover)

# environmental exposures:
ozone <- read_csv("outputs/comparison/ozone_2026-06-01.csv")

pm25 <- read_csv("outputs/comparison/pm25_2026-06-01.csv")

pesticides <- read_csv("outputs/comparison/pesticides_2026-06-02.csv")

traffic <- read_csv("outputs/comparison/traffic_proximity_2026-06-26.csv")

# environmental effects: 
npl_prox <- read_csv("outputs/comparison/npl_proximity_2026-06-01.csv")

rmp_prox <- read_csv("outputs/comparison/rmp_proximity_2026-06-01.csv")

haz_prox <- read_csv("outputs/comparison/hazardous_waste_proximity_2026-06-01.csv")

# urban vs. rural 
ca_urban_areas <- read_sf("data/ca_sample/tl_2020_us_uac20/tl_2020_us_uac20.shp") %>% 
  filter(str_ends(NAME20, " CA")) 
```

Next, let's combine into one data frame and save outputs. Again thinned subsets can be retrieved via the function `load_ca_samples` found in the `analysis/` folder

```{r}
# combine into single data frame 
final_df <- list(ca_residences,
                 flood_risk, wildfire_risk, heat_exp, 
                 canopy_cover, ozone, pm25, pesticides, 
                 traffic,npl_prox, rmp_prox, haz_prox) %>% 
  purrr::reduce(left_join, by = "object_id") %>% 
  #make object id a character
  mutate(object_id = as.character(object_id)) 

# add urban vs rural to entire data set
ca_urban_areas <- st_transform(ca_urban_areas,crs = crs(final_df))
final_df$urban <- lengths(st_intersects(final_df,ca_urban_areas))>0 

#save as csv and as shapefile for future analysis
# entire sample:
final_df %>% st_as_sf() %>% write_sf(paste0("outputs/comparison/final_df_all_", Sys.Date(), ".gpkg"))

final_df %>%
  select(-geom) %>% 
  write_csv(paste0("outputs/comparison/final_df_all_", Sys.Date(), ".csv"))
```
