---
title: "Code to process all indicators and create final data frame for sampled California residential areas"
author: "Jackson Phippen"
date: "`r Sys.Date()`"
output: 
  html_document: 
    self_contained: false
---

## Process Indicators

Currently (as of June 1st, 2026), there are three subsets of the same sample:

1.  **All 5,000** sampled points via population weighting
2.  Thinned sampled points to be at least **1km** from all prisons and neighboring points
3.  Thinned sampled points to be at least **5km** from all prisons and neighboring points

To avoid weird computing, I will compute indicators for all 5,000 points and correctly subset points for future analysis.

There are 11 indicators in this code base. The majority of which are calculated in R, and have their own functions in the `R/` folder. Two indicators (canopy cover and heat exposure) are calculated via Google Earth Engine and have stand alone Python scripts in the 'python/' folder. 

Component scores will not be calculated for these data. 

First set up the environment with all required libraries via "setup.R" file, and load in data of interest. 

```{r}
# setup environment
source("setup.R")

# read in sampled points and california prisons
ca_residences <- read_sf("data/ca_sample/all_ca_residences.gpkg")
thin_1km <- read_sf("data/ca_sample/1km_ca_residences.gpkg")
thin_5km <- read_sf("data/ca_sample/5km_ca_residences.gpkg")

ca_prisons <- read_sf("data/phase2/processed/prisons/study_prisons.gpkg") %>% 
  filter(STATE == "CA")
```


#### Flood Risk

```{r}
# flood risk (takes 1-2 days to run) (Last run time was 1.72 hours, not sure how/why)
flood_risk <- calc_flood_risk(
  sf_obj = ca_residences,
  dist = 1000,
  id_column = "object_id",
  save = TRUE,
  out_path = "outputs/comparison/"
)
```

#### Wildfire Risk

```{r}
# wildfire risk
wildfire_risk <- calc_wildfire_risk(
  sf_obj = ca_residences,
  file = "data/phase2/raw/wildfire_risk/Data/whp2023_GeoTIF/",
  dist = 1000,
  conus_only = TRUE,
  id_column = "object_id",
  save = TRUE,
  out_path = "outputs/comparison/"
)
```

#### Ozone

```{r}
# ozone
ozone <- calc_ozone(
  sf_obj = ca_residences, 
  years    = c(2019, 2021, 2022),
  folder   = "data/phase2/raw/ozone",
  id_column = "object_id",
  save     = TRUE,
  out_path = "outputs/comparison/"
)
```

#### PM 2.5

```{r}
# pm2.5 (may change dataset, check for more recent years)
pm25 <-
  calc_pm25(
    sf_obj = ca_residences,
    folder = "data/phase2/processed/PM2.5",
    dist = 1000,
    years = c(2021, 2023),
    id_column = "object_id",
    save = TRUE,
    out_path = "outputs/comparison/"
  )
```

#### Pesticides

```{r}
pesticides <- calc_pesticides(
  ca_residences,
  folder = "data/phase2/raw/pesticide/PEST-CHEMGRIDS_v2/NC/",
  dist = 0, # resolution is 5km
  id_column = "object_id",
  save = TRUE,
  out_path = "outputs/comparison/")
```

#### Traffic Proximity

```{r}
#takes 1.5 days to run
traffic_prox <- calc_traffic_proximity(
  sf_obj = ca_residences,
  file = "data/phase2/processed/traffic/aadt_2023.RData",
  dist = 500,
  id_column = "object_id",
  save = TRUE,
  out_path = "outputs/comparison/"
)
```

#### Risk Management Plan (RMP) Facility Proximity

```{r}
#calculate Risk Management Plan (RMP) facility proximity
rmp_prox <- calc_rmp_proximity(
  sf_obj = ca_residences,
  file = "data/phase2/raw/RMP/EPA RMP Spreadsheets — The Data Liberation Project - Facilities.csv",
  dist = 5000,
  id_column = "object_id",
  save = TRUE,
  out_path = "outputs/comparison/"
)
```

#### NPL/Superfund Facility Proximity

```{r}
# calculate NPL facility proximity
npl_prox <- calc_npl_proximity(
  sf_obj = ca_residences,
  file = "data/phase2/processed/NPL/NPL_arc_geocode.shp",
  dist = 5000,
  id_column = "object_id",
  save = TRUE,
  out_path = "outputs/comparison/"
)
```

#### Hazardous Waste Facility Proximity

```{r}
# calculate Haz waste facility proximity
haz_prox <- calc_haz_waste_proximity(
  sf_obj = ca_residences,
  file = "data/phase2/processed/hazardous_waste/TSD_LQGs.csv",
  dist = 5000,
  id_column = "object_id",
  save = TRUE,
  out_path  = "outputs/comparison/"
)
```
