---
title: "Code to process all indicators and create final data frame for ICE facilities"
author: "Caitlin C. Mothes, PhD"
date: "`r Sys.Date()`"
output: html_document
---

## Process Indicators

There are a total of 11 indicators in this code base. The majority of them are calculated in R and each have their own function (found in the 'R/' folder). Two indicators, heat exposure and canopy cover, were calculated using Google Earth Engine and each have a stand alone Python script in the 'python/' folder.

This workflow is if you want to calculate the indicators individually. If you want to calculate indicators as their component scores, you can do so in the 'process_components.Rmd' file.

First set up the environment with all the required libraries (using the 'setup.R' script to install and load packages) and the spatial boundaries (in `sf` format) of interest. In this workflow we are calculating these indicators for state and federal prisons, but **these functions will work on any sf polygon file in the U.S.**

```{r}
# set up environment -----------------
source("setup.R")

# read in processed ICE facilities
ice_facilities <- read_sf("data/ice/ice_detention_facilities.gpkg")
```

#### Flood Risk

```{r}
# flood risk (takes 1-2 days to run)
flood_risk <- calc_flood_risk(
  sf_obj = ice_facilities,
  dist = 1000,
  id_column = "object_id",
  save = TRUE,
  out_path = "outputs/ice/"
)

```

#### Wildfire Risk

```{r}
# wildfire risk
wildfire_risk <- calc_wildfire_risk(
  sf_obj = ice_facilities,
  file = "data/phase2/raw/wildfire_risk/Data/whp2023_GeoTIF/",
  dist = 1000,
  conus_only = TRUE,
  id_column = "object_id",
  save = TRUE,
  out_path = "outputs/ice/"
)
```

#### Ozone

```{r}
# ozone
ozone <- calc_ozone(
  sf_obj = ice_facilities, 
  years    = c(2019, 2021, 2022),
  data_dir = "data/phase2/raw/ozone",
  id_column = "object_id",
  save     = TRUE,
  out_path = "outputs/ice/"
)
```

#### PM 2.5

```{r}
# pm2.5 (may change dataset, check for more recent years)
pm25 <-
  calc_pm25(
    sf_obj = ice_facilities,
    folder = "data/phase2/processed/PM2.5",
    dist = 1000,
    years = c(2021, 2023),
    id_column = "object_id",
    save = TRUE,
    out_path = "outputs/ice/"
  )
```

#### Pesticides

```{r}
pesticides <- calc_pesticides(prisons,
                              folder = "data/phase2/raw/pesticide/PEST-CHEMGRIDS_v2/NC/",
                              dist = 0, # resolution is 5km
                              id_column = "object_id",
                              save = TRUE,
                              out_path = "outputs/ice/")
```

#### Traffic Proximity

```{r}
# traffic proximity (takes 1.5 days to run on Desktop comp)
traffic_prox <- calc_traffic_proximity(
  sf_obj = ice_facilities,
  file = "data/phase2/processed/traffic/aadt_2023.RData",
  dist = 500,
  id_column = "object_id",
  save = TRUE,
  out_path = "outputs/ice/"
)
```

#### Risk Management Plan (RMP) Facility Proximity

```{r}
# calculate Risk Management Plan (RMP) facility proximity
rmp_prox <- calc_rmp_proximity(
  sf_obj = ice_facilities,
  file = "data/phase2/raw/RMP/EPA RMP Spreadsheets — The Data Liberation Project - Facilities.csv",
  dist = 5000,
  id_column = "object_id",
  save = TRUE,
  out_path = "outputs/ice/"
)

```

#### NPL/Superfund Facility Proximity

```{r}
# calculate NPL facility proximity
npl_prox <- calc_npl_proximity(
  sf_obj = ice_facilities,
  file = "data/phase2/processed/NPL/NPL_arc_geocode.shp",
  dist = 5000,
  id_column = "object_id",
  save = TRUE,
  out_path = "outputs/ice/"
)
```

#### Hazardous Waste Facility Proximity

```{r}
# calculate Haz waste facility proximity
haz_prox <- calc_haz_waste_proximity(
  sf_obj = ice_facilities,
  file = "data/phase2/processed/hazardous_waste/TSD_LQGs.csv",
  dist = 5000,
  id_column = "object_id",
  save = TRUE,
  out_path  = "outputs/ice/"
)

```

### Component Score Calculation

Now to calculate component scores, you can run the following chunks of code. In this example I read in the output datasets from running each of the functions above. Since some of the functions take a long time to run, you may not create every object in a single session anyways.

#### Climate Scores

Note that heat exposure and canopy cover were saved in the 'data/processed' folder and exported from GEE instead of the 'outputs/' folder

```{r}
## climate scores -------------

flood_risk <- read_csv("outputs/ice/flood_risk_2026-05-20.csv") %>%
  select(object_id, flood_risk = flood_risk_percent)

wildfire_risk <- read_csv("outputs/ice/wildfire_risk_2026-05-20.csv")

heat_exp <- read_csv("data/ice/ice_lst_daily_MODIS_2026-05-19.csv") %>% 
  group_by(object_id) %>%
  summarise(lst_avg = median(LST_mean, na.rm = TRUE))

canopy_cover <- read_csv("data/ice/ice_canopy_2026-05-19.csv") %>%
  rename(percent_tree_cover = NLCD_Percent_Tree_Canopy_Cover)


climate_scores <- list(flood_risk, wildfire_risk, heat_exp, canopy_cover) %>%
  # convert FACILITYID to character for all to make sure they join
  purrr::map(~ .x %>% mutate(object_id = as.character(object_id))) %>% 
  purrr::reduce(left_join, by = "object_id") %>%
  # calculate percentile columns for each raw indicator
  mutate(across(
    where(is.numeric),
    .fns = list(pcntl = ~ cume_dist(.) * 100),
    .names = "{col}_{fn}"
  )) %>%
  # need to inverse canopy cover since high value is good
  mutate(percent_tree_cover_pcntl = cume_dist(desc(percent_tree_cover)) * 100) %>%
  rowwise() %>%
  # calculate climate component score (average all indicator percentile values per prison
  mutate(climate_score = gm_mean(c_across(contains("pcntl"))))
```

#### Environmental Exposures Scores

```{r}
ozone <- read_csv("outputs/ice/ozone_2026-05-20.csv") %>% select(object_id, mean_ozone)

pm25 <- read_csv("outputs/ice/pm25_2026-05-20.csv")

pesticides <- read_csv("outputs/ice/pesticides_2026-05-20.csv")

traffic <- read_csv("outputs/ice/traffic_proximity_2026-05-20.csv")

exposure_scores <- list(ozone, pm25, pesticides, traffic) %>%
  # convert FACILITYID to character for all to make sure they join
  purrr::map(~ .x %>% mutate(object_id = as.character(object_id))) %>% 
  purrr::reduce(left_join, by = "object_id") %>%
  # calculate percentile columns for each raw indicator
  dplyr::mutate(across(
    where(is.numeric),
    .fns = list(pcntl = ~ cume_dist(.) * 100),
    .names = "{col}_{fn}"
  )) %>%
  rowwise() %>%
  # calculate climate component score (geometric mean of indicator percentiles)
  mutate(exposure_score = gm_mean(c_across(contains("pcntl"))))


```

#### Environmental Effects Scores

```{r}
npl <- read_csv("outputs/ice/npl_proximity_2026-05-21.csv")

rmp <- read_csv("outputs/ice/rmp_proximity_2026-05-21.csv")

haz <- read_csv("outputs/ice/hazardous_waste_proximity_2026-05-21.csv")

effects_scores <- list(npl, rmp, haz) %>%
  purrr::reduce(left_join, by = "object_id") %>%
  # make facility ID character
  mutate(object_id = as.character(object_id)) %>%
  # calculate percentile columns for each raw indicator
  dplyr::mutate(across(
    where(is.numeric),
    .fns = list(pcntl = ~ cume_dist(.) * 100),
    .names = "{col}_{fn}"
  )) %>%
  rowwise() %>%
  # calculate climate component score (average all indicator percentile values per prison
  mutate(effects_score = gm_mean(c_across(contains("pcntl"))))
```

### Final Data Frame

Now using the component scores data frames we can create a single, final data frame with an overall vulnerability index.

```{r}
final_df <- list(climate_scores, exposure_scores, effects_scores) %>%
  purrr::reduce(left_join, by = "object_id") %>%
  # remove rowwise
  ungroup() %>%
  mutate(
    final_risk_score = rowMeans(select(., contains("score"))),
    final_risk_score_pcntl = cume_dist(final_risk_score) * 100
  ) %>% 
  # join with original ice facility metadata
  mutate(object_id = as.character(object_id)) %>% 
  left_join(ice_facilities %>% mutate(object_id = as.character(object_id)), by = "object_id")

# save as shapefile and csv
final_df %>% st_as_sf() %>% write_sf(paste0("outputs/ice/final_df_", Sys.Date(), ".gpkg"))

final_df %>% 
  select(-geom) %>% 
  write_csv(paste0("outputs/ice/final_df_", Sys.Date(), ".csv"))
```
