---
title: "exploratory_data_analysis"
output: html_document
---

```{r}
#source packages
source("setup.R")

library(DataExplorer) #quickly export images
```

First, let us load in the applicable data

```{r}
#for now, use the final csv from phase 1 of the project
data <- read.csv("outputs/final_df_2023-08-31.csv") %>% data.frame()

#a precursory view of what we have. 
#summary(data)

```

Next, it is desirable to have a sneak peak of the indicators, let's first just run the indicators through summary

```{r}
#select indicators
indicator_data <- data %>% 
  select(flood_risk, 
         wildfire_risk,
         lst_avg,
         percent_tree_cover,
         mean_ozone, #will need to remove for new analysis (for now)
         avg_pm25,
         pesticides,
         trafficProx,
         npl_prox,
         rmp_prox,
         haz_prox) 

#quick 5-number summary of our data. 
#not sure what is happening with some of our proximity measures, will need to look a bit closer.
summary(indicator_data)


```

Ok, things here seem promising, let's take a look at the inf's.

When computing any of our proximity scores, we are using a "Weighted Proximity Score", which ensures that features closest to the value of interest "count more" than features further away. This score is expressed as\
$$
\text{Proximity Score} = \sum_{i=1}^N \frac{1}{d_i}, i= 1, 2, \ldots, N
$$ 
where $d_i$ is the distance with in a specific boundary, or in our case nearest if there are no features with in the boundary. Note, if a score comes up as "Inf," this is `r`'s way of rounding a value $d_i$ that is very close to/is $0$. **What was done previously when working with "Inf?"** (Set value to 0.)

To get a glimpse of what's going on, let's map the prisons where an inf is happening in Traffic Proximity and Risk Management Plan Proximity. 

```{r}
#snag the prisons that are causing issues.
infs <- data %>% filter(is.infinite(trafficProx)|is.infinite(rmp_prox))
#head(infs)

#recover the boundaries to plot
old_prisons <- read_sf("data/phase1/processed/prisons/study_prisons.shp") 
flagged_prisons <- old_prisons %>% filter(FACILITYID%in%infs$FACILITYID)

#let's handle traffic proximity first
#load in traffic data
load("data/phase1/processed/traffic_proximity/aadt_2018.RData")
traffic_inf <- flagged_prisons %>% filter(FACILITYID == "10003116")
#step through what calc traffic is doing. First add buffer
buffer <- st_buffer(traffic_inf,500)

#find roads near this point
roads_nearby <- st_crop(aadt_2018,buffer) %>% 
  mutate(distance=as.numeric(st_distance(.,traffic_inf))) %>%
  mutate(Overlap = distance == 0)
           
ggplot()+
  geom_sf(data=roads_nearby,aes(color=Overlap))+
  geom_sf(data=buffer,fill=NA,color="blue",linetype="dashed")+
  geom_sf(data=traffic_inf,color="red",size=3)+
  theme_minimal()+
  labs(
    title="Flagged Prison with 'Inf' Traffic Proximity"
  )


```


Let's repeat this for rmp_proximity, too
```{r}
rmp_inf <- flagged_prisons[2,]
rmp_prox <- read.csv("data/phase1/raw/EPA_RMP/EPA_Emergency_Response_(ER)_Risk_Management_Plan_(RMP)_Facilities.csv") %>% 
  st_as_sf(coords = c("LONGITUDE8", "LATITUDE83"), crs = 4269) %>%
  st_transform(crs = st_crs(rmp_inf))

#find distances
rmp_nearby <- rmp_inf %>%
    mutate(find_points = st_is_within_distance(., rmp_prox, dist = 500)) %>%
    unnest(find_points)

#snag the county the prison is in
library(tigris)
mn_counties <- counties(state="MN")
pine_county <- mn_counties %>% filter(NAME == "Pine")

tm_shape(pine_county)+
  tm_lines()+
tm_shape(rmp_inf)+
  tm_polygons()+
  tm_shape(rmp_prox[rmp_nearby$find_points,])+
  tm_dots(size=1)

ggplot() +
  geom_sf(data = rmp_inf, fill = NA, color = "black") +
  geom_sf(
    data = rmp_prox[rmp_nearby$find_points, ],
    color = "blue",
    size = 2
  ) +
  theme_minimal()+
  labs(
    title="Flagged Prison with 'Inf' rmp_proximity"
  )

```

```{r,eval = FALSE}

#let's export some uni-dimensional stats for what we have
DataExplorer::create_report(indicator_data,
                            config=list(
                              introduce=list(),
                              plot_missing=list(),
                              plot_histogram=list(),
                              plot_boxplot=list(),
                              plot_correlation=list(),
                              plot_qq=list(),
                              plot_scatterplot=list()
                            ))
```




```{r}
#ok. we need to speed this up and consider data not yet in the final data set form. 

#these data sets are FACILITY ID and INDICATOR Data, so column two is of interest, primarily. 
library(patchwork)
analyze_data <- function(file){
  #load in data
  data <- read.csv(paste0("outputs/",file)) %>% data.frame()
  
  #display histogram, and box plot, and summarize
  fivenumsum <- summary(data[,2])
  
  hist <- data %>% ggplot(aes(x=data[,2]))+
    geom_histogram()+
    theme_minimal()
  
  box <- data %>% ggplot(aes(x=data[,2]))+
    geom_boxplot()+
    coord_flip()+
    theme_minimal()
  
  plot <- hist +box
  
  outputs <- list(fivenumsum,plot)
  return(outputs)
}

new_files <- list("hazardous_waste_proximity_2026-02-12.csv",
                  "npl_proximity_2026-03-11.csv",
                  "pm25_2026-03-26.csv",
                  "rmp_proximity_2026-02-11.csv",
                  "wildfire_risk_2026-02-12.csv")

summaries <- list()
plots <- list()
for (i in 1:length(new_files)){
  print(paste0(new_files[i]))
  temp <- analyze_data(new_files[i])
  
  summaries[i] <- temp[1]
  plots[i] <- temp[2]
  i+1
}

```
```{r}
new_prisons <- read_sf("data/phase2/processed/prisons/study_prisons.gpkg")

haz_prox <- read.csv("outputs/hazardous_waste_proximity_2026-02-12.csv") %>% 
  data.frame()

haz_data <- read.csv("data/phase2/processed/hazardous_waste/TSD_LQGs.csv") %>% 
  filter(!is.na(LONGITUDE83) | !is.na(LATITUDE83)) %>%
    st_as_sf(coords = c("LONGITUDE83", "LATITUDE83"), crs = 4269) %>%
    st_transform(crs = st_crs(new_prisons))

summary(haz_prox)

infs <- haz_prox %>% filter(haz_prox=="Inf")

prison_infs <- new_prisons %>% filter(FACILITYID%in%infs$FACILITYID)

dist <- 5000

prison_dist <- prison_infs %>%
  mutate(find_points = st_is_within_distance(., haz_data, dist = dist))

first <- prison_dist[1,]
mapview(prison_infs[1,])+
  mapview(haz_data[prison_dist[1,28],])

mapview(prison_infs[2,])+
  mapview(haz_data[prison_dist[2,28],])

mapview(prison_infs[3,])+
  mapview(haz_data[prison_dist[3,28],])

mapview(prison_infs[4,])+
  mapview(haz_data[prison_dist[4,28],])
```
```{r}
npl_data <- read_sf("data/phase2/processed/NPL/NPL_arc_geocode.shp")

npl_prox <- read.csv("outputs/npl_proximity_2026-03-11.csv")

summary(npl_prox) #at first glance, the doesn't seem to be anything sus here. Carry on brother
```
```{r}
pm25_prox <- read.csv("outputs/pm25_2026-03-26.csv")

summary(pm25_prox) #also seems fine. but like who knows
```
```{r}
rmp_prox <- read.csv("outputs/rmp_proximity_2026-02-11.csv")
rmp_data <- read.csv("data/phase2/raw/RMP/EPA RMP Spreadsheets — The Data Liberation Project - Facilities.csv")%>%
    st_as_sf(coords = c("Lng", "Lat"), crs = 4269) %>%
    st_transform(crs = st_crs(new_prisons))

summary(rmp_prox) #ah, caught some some infs here. 

rmp_infs <- rmp_prox %>% filter(rmp_prox=="Inf")

rmp_prison_inf <- new_prisons %>% filter(FACILITYID%in%rmp_infs$FACILITYID) #oh, this may be different than what we pulled in the old data set... weird!

#find points: 2353,9495
#st_is_within_distance(rmp_prison_inf, rmp_data,5000)

mapview(rmp_prison_inf)+
  mapview(rmp_data[c(2353,9495),]) #looks like second point over laps
```

```{r}
wildfire_risk <- read.csv("outputs/wildfire_risk_2026-02-12.csv")

summary(wildfire_risk) #how is this interpreted/what does a max of 39,254.352 mean???

#how many of these are just absolutely packing?
big_risk <- wildfire_risk %>% filter(wildfire_risk>1000)
lil_risk <- wildfire_risk %>% filter(wildfire_risk<1000)

hist(lil_risk$wildfire_risk)
boxplot(lil_risk$wildfire_risk)
summary(whp_values)
quantile(wildfire_risk, probs = c(0, .25, .5, .75, .9, .99, 1), na.rm = TRUE)
hist(wildfire_risk$wildfire_risk, breaks = 50)
```

