---
title: "process_prisons"
output: html_document
---

## Process raw prison dataset to study sites

```{r}
source("setup.R")
```

### Read in HIFLD dataset

*Last updated Aug 2025*

```{r}
prisons <- read_sf("data/phase2/raw/prisons/Prison_Boundaries.shp") %>% 
  #filter out just state and federal
  filter(TYPE %in% c("STATE", "FEDERAL")) %>% 
  #filter just U.S. (not territories)
  filter(COUNTRY == "USA") %>% 
  # filter out prisons with 0 or NA population and that are designated as "closed"
  #filter(POPULATION > 0) %>% no longer using population as a filter
  filter(STATUS == "OPEN") %>% 
  # some invalid geometries, fix that
  st_make_valid()

# save as geopackage
write_sf(prisons, "data/phase2/processed/prisons/study_prisons.gpkg")

# Create centroids and save as csv with lat/long for data viz purposes
prisons %>% 
  st_centroid() %>% 
  mutate(long = unlist(map(.$geometry,1)),
         lat = unlist(map(.$geometry,2))) %>%
  st_drop_geometry() %>% 
  select(FACILITYID, NAME, long, lat) %>% 
  write_csv("data/phase2/processed/prisons/prison_centroids.csv")
```

------------------------------------------------------------------------

## Comparisons with previous version:

2024 dataset used in Phase 1:

```{r}
prisons_old <- read_sf("data/phase1/raw/Prisons/Prison_Boundaries.shp") %>% 
  #filter out just state and federal
  filter(TYPE %in% c("STATE", "FEDERAL")) %>% 
  #filter just U.S. (not territories)
  filter(COUNTRY == "USA") %>% 
  # filter out prisons with 0 or NA population and that are designated as "closed"
  filter(POPULATION > 0) %>% 
  filter(STATUS == "OPEN")
```

There are a lot less prisons in the updated dataset using our previous filter options. View the mismatches:

```{r}
prison_add <- prisons %>% filter(!FACILITYID %in% prisons_old$FACILITYID)

prison_drop <- prisons_old %>% filter(!FACILITYID %in% prisons$FACILITYID)
```

Explore prisons dropped:

```{r}
# any pattern in states
prison_drop %>% 
  st_drop_geometry() %>% 
  group_by(STATE) %>% 
  count() %>% 
  arrange(-n)


# any pattern in juristiction
prison_drop %>% 
  st_drop_geometry() %>% 
  group_by(TYPE) %>% 
  count() %>% 
  arrange(-n)

# almost all state prisons...
```

See if dropped prisons are in the new dataset at all and view comparisons

```{r}
prisons_new_raw <- read_sf("data/phase2/raw/prisons/Prison_Boundaries.shp")

x <- prisons_new_raw %>% 
  filter(FACILITYID %in% prison_drop$FACILITYID)


```

Ok, main difference is these prisons are still categorized as OPEN but the population data went from a value to -999. Move forward disregarding the POPULATION column due to high uncertainty.
