---
title: "process_ca_sample"
author: "Jackson Phippen"
date: "`r Sys.Date()`"
output: html_document
---

This workflow processes the high-population density raster from Meta filtered for the state of California and generates samples population-weighted samples and considers various thinning strategies. 

Setup for workflow: 
```{r}
source("setup.R")
library(FNN) #for quick nearest neighbor calculations
library(mapview) #quick interactive plots

#read in prisons and raster
ca_prisons <- read_sf("data/phase2/processed/prisons/study_prisons.gpkg") %>% 
  filter(STATE == "CA")

ca_pop_raster <- rast("data/phase2/raw/population/ca_30m_pop_density_hrsl_2020.tif")
```

The raster has a resolution of ~30m for each cell. For processing, let's aggregate w/ factor 10, round to the nearest integer, and drop all zeros.
```{r}
pop_aggregate <- ca_pop_raster %>% 
  terra::aggregate(fact = 10, 
                   fun = sum, 
                   na.rm = TRUE)

#next, let's round to the nearest integer
pop_aggregate <- terra::round(pop_aggregate)

#now, let's drop all 0's. 
pop_aggregate[pop_aggregate==0] <- NA
```

Next, let's sample points with respect to population. i.e. cells with larger population values will be drawn more frequently. For reproducibility, let's set our seed. 
```{r}
set.seed(513)
#to speed up calculations, lets first convert values to a data frame
pop_vals <- as.data.frame(pop_aggregate,cells=TRUE, na.rm = TRUE)
colnames(pop_vals) <- c("cell", "pop")

#add a weight/probability column
pop_vals <- pop_vals %>% 
  filter(pop>0) %>% #remove zeros if values weren't changed to NA's
  mutate(prob = pop/sum(pop))

#now let's attempt to sample points without replacement
sampled_points <- sample(
  x = pop_vals$cell,
  size = 5000,
  replace = FALSE, 
  prob = pop_vals$prob
)

# keep sampled rows INCLUDING population values
sampled_df <- pop_vals %>%
  filter(cell %in% sampled_points)

# add coordinates
coords <- xyFromCell(pop_aggregate, sampled_df$cell)
sampled_df <- sampled_df %>%
  mutate(object_id = 1:5000) %>% 
  mutate(
    lon = coords[,1],
    lat = coords[,2]
  )

# convert to sf object
sampled_sf <- st_as_sf(
  sampled_df,
  coords = c("lon", "lat"),
  crs = crs(pop_aggregate)
)

#store points for thinning and nearest neighbor analysis
sampled_pts <- sampled_sf

#reproject the points to meters
sampled_pts <- st_transform(sampled_pts,crs=3310)
ca_prisons <- ca_prisons %>% st_transform(crs=3310)
```

Next, let's thin points that are 1km and 5km apart. Our spatial thinning algorithm with exclusion constraint works as follows:

This function performs a randomized greedy thinning of point data
subject to two constraints:
  (1) Minimum separation between retained points (threshold)
  (2) Minimum distance from any prison location (threshold)

Algorithm steps:
  1. Ensure both datasets (pts and prisons) share a common CRS.
  2. Extract coordinate matrices for points and prisons.
  3. Randomly permute the order of candidate points to avoid
     ordering bias in selection.
  4. Initialize the kept set with the first randomly selected point.
  5. Iteratively evaluate each remaining point:
       a. Compute Euclidean distance to all already-kept points.
       b. Compute Euclidean distance to all prison locations.
       c. Retain the point only if it is at least `threshold`
          distance from BOTH:
             - all previously kept points
             - all prisons
  6. Return the subset of points that satisfy both constraints.

Output:
  A spatial subset of `pts` satisfying minimum spacing and
  minimum distance-from-prison constraints under a greedy,
  order-randomized selection procedure.

```{r}

thin_function <- function(pts, prisons, threshold, seed = 123){
  #set a seed for reproducibility
  set.seed(seed = seed)
  
  if(!st_crs(pts)==st_crs(prisons)){
    prisons <- st_transform(prisons, crs = st_crs(pts)) #should be Albers for comparison... 
  }
  
  #convert coords to a coordinate matrix
  coords <- st_coordinates(pts)
  prison_coords <- st_coordinates(prisons) 
  
  #enforce an order 
  n <- nrow(pts)
  order <- sample(1:n)
  
  #initialize
  keep <- rep(FALSE, n)
  keep[order[1]] <- TRUE # we always keep the first point
  kept_coords <- coords[order[1],,drop=FALSE]
  
  #loop over remaining points
  for (i in 2:n){
    idx <- order[i]
    candidate_pt <- coords[idx,,drop=FALSE]
    
    #compute euclidean distance to kept coordinates
    d_pts <- sqrt(
      (kept_coords[,1] - candidate_pt[1,1])^2 +
        (kept_coords[,2] - candidate_pt[1,2])^2
    )
    
    d_prisons <- sqrt(
      (prison_coords[,1]-candidate_pt[,1])^2+
        (prison_coords[,2]-candidate_pt[,2])^2
    )
    #check threshold
    if (min(d_pts)>=threshold&& min(d_prisons)>=threshold){
      keep[idx] <- TRUE
      kept_coords <- rbind(kept_coords,candidate_pt)
    }
    
  }
  return(pts[keep,])
}
```

Return thinned points for 1 km, and 5 km. 
```{r}
thin_1km <- thin_function(sampled_pts,ca_prisons, 1000,513)
thin_5km <- thin_function(sampled_pts,ca_prisons, 5000,513)
```

For further analysis, let's add a buffer with radius 140.4 m to match the median square footage for Californian prisons. 
```{r}
#add buffers
all_points <- st_buffer(sampled_pts, 140.4) #includes intersections with self and prisons
thin_1km_buffers <- st_buffer(thin_1km, 140.4)
thin_5km_buffers <- st_buffer(thin_5km, 140.4)
```

Finally, let's save these polygons. 
```{r}
#all points
write_sf(all_points,"data/ca_sample/all_ca_residences.gpkg")
#1km thin
write_sf(thin_1km_buffers,"data/ca_sample/1km_ca_residences.gpkg")
#5km thin
write_sf(thin_5km_buffers,"data/ca_sample/5km_ca_residences.gpkg")
```

