---
title: "ICE Detention Facilities: Data Pull & Review"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Setup

```{r packages}
library(sf)
library(tidyverse)
library(janitor)
library(mapview)
library(httr2)
library(tigris)
```

## Pull Data from FeatureServer

This layer has a max record count of 1,000 per request, so we paginate
using `resultOffset` until all features are retrieved.

```{r pull-data}
base_url <- "https://services1.arcgis.com/SIYkiqjmENweC50g/ArcGIS/rest/services/rolling_updated_feb12/FeatureServer/0/query"

ice_facilities <- st_read(paste0(
    base_url,
    "?where=1%3D1",
    "&outFields=*",
    "&returnGeometry=true",
    "&outSR=4326",
    "&f=geojson")
)

message("Total features retrieved: ", nrow(ice_facilities))
```

## Initial Review

### Dimensions & Field Names

```{r review-dims}
dim(ice_facilities)
names(ice_facilities)
```

### Field Types

```{r review-types}
glimpse(ice_facilities)
```

### Missing Values by Field

```{r review-na}
ice_facilities |>
  st_drop_geometry() |>
  summarise(across(everything(), ~ sum(is.na(.)))) |>
  pivot_longer(everything(), names_to = "field", values_to = "n_missing") |>
  mutate(pct_missing = round(n_missing / nrow(ice_facilities) * 100, 1)) |>
  arrange(desc(n_missing))
```

### Detainment Summary Statistics

```{r review-detainment}
ice_facilities |>
  st_drop_geometry() |>
  select(Total_Detainment, Total_Female_Detainment, Total_Male_Detainment, Total_Non_Criminal, FY26_ALOS) |>
  summary()
```

### Facility Type Breakdown

```{r review-types-table}
ice_facilities |>
  st_drop_geometry() |>
  count(Type_Detailed, sort = TRUE)
```

### Last Updated Distribution

```{r review-dates}
ice_facilities |>
  st_drop_geometry() |>
  mutate(last_updated = as.POSIXct(last_updated / 1000, origin = "1970-01-01", tz = "UTC") |> as.Date()) |>
  count(last_updated, sort = TRUE)
```

### Geometry Check

```{r review-geometry}
# CRS
st_crs(ice_facilities)

# Any invalid geometries?
any(!st_is_valid(ice_facilities))

# Any empty geometries?
any(st_is_empty(ice_facilities))

# Bounding box sanity check (should be roughly CONUS + territories)
st_bbox(ice_facilities)
```

### Interactive Map

```{r review-map}
mapview(
  ice_facilities,
  zcol       = "Total_Detainment",
  layer.name = "Total Detainment",
  cex        = 4
)
```

### Coordinate Outliers

Points outside US bounding box (Puerto Rico and Guam)

```{r review-coords}
coords <- ice_facilities |>
  st_coordinates() |>
  as_tibble() |>
  rename(lon = X, lat = Y)

ice_facilities |>
  st_drop_geometry() |>
  bind_cols(coords) |>
  filter(lon < -180 | lon > -60 | lat < 15 | lat > 72) |>
  select(Name, Full_Address, lon, lat)
```



## Clean & Standardize

```{r clean}
# need to filter to CONUS
conus <- states(cb = TRUE) |>
  filter(!STUSPS %in% c("AK", "HI", "PR", "GU", "VI", "MP", "AS")) |>
  st_union() |>
  st_transform(st_crs(ice_facilities))

ice_clean <- ice_facilities |>
  st_filter(conus) |>
  clean_names() |>                          # snake_case field names
  mutate(
    last_updated = as.POSIXct(last_updated / 1000, origin = "1970-01-01", tz = "UTC") |> as.Date()) |>
  # Reproject to EPSG:5070 (Albers Equal Area) — standard for this project
  st_transform(5070)
```


## Save to data/ice/

```{r save}
out_dir <- here::here("data", "ice")
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

out_path <- file.path(out_dir, "ice_detention_facilities.gpkg")

st_write(ice_clean, out_path, delete_dsn = TRUE, append = FALSE)

message("Saved: ", out_path)
message("Features: ", nrow(ice_clean))
message("CRS: ", st_crs(ice_clean)$input)

# also save as csv for GEE upload
ice_clean %>%
  mutate(wkt = st_as_text(geometry)) %>% # use 'wkt' as geometry info for GEE upload
  st_drop_geometry() %>%
  write_csv("data/ice/ice_detention_centers.csv")
```

