---
title: "U.S. State & Federal Prisons — Environmental Vulnerability Index: Exploration & QA"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
author: "Caitlin Mothes, PhD"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    number_sections: true
    code_folding: hide
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = TRUE,
  warning    = FALSE,
  message    = FALSE,
  fig.width  = 9,
  fig.height = 5,
  dpi        = 150,
  out.width  = "100%"
)
```

```{r packages}
library(tidyverse)
library(sf)
library(scales)
library(knitr)
library(kableExtra)
library(patchwork)
library(janitor)
library(ggridges)
library(leaflet)
library(leaflet.extras)
library(glue)
```

------------------------------------------------------------------------

# Overview

This report provides a structured QA and exploratory analysis of the U.S. state and federal prison Environmental Vulnerability Index dataset. It is organized in three parts:

1.  **Data inventory & flags** — completeness, range checks, structural anomalies
2.  **Indicator-level exploration** — distributions and outliers for each raw indicator
3.  **Component & index summaries** — score distributions, facility rankings, cross-component patterns

```{r load-data}
# ── Load data ──────────────────────────────────────────────────────────────────
df <- read_csv("outputs/final_df_2026-04-29.csv", show_col_types = FALSE) %>%
  mutate(
    facility_type = factor(TYPE),
    secure_level  = factor(SECURELVL),
    full_address  = glue::glue("{ADDRESS}, {CITY}, {STATE} {ZIP}")
  )

df_sf <- read_sf("outputs/final_df_2026-04-29.gpkg")

n_facilities <- nrow(df)

# ── Column groupings ───────────────────────────────────────────────────────────
raw_climate  <- c("flood_risk", "wildfire_risk", "lst_avg", "percent_tree_cover")
raw_exposure <- c("mean_ozone", "avg_pm25",
                  "pesticide_mean_kg", "trafficProx")
raw_effects  <- c("npl_prox", "rmp_prox", "haz_prox")
raw_all      <- c(raw_climate, raw_exposure, raw_effects)

pcntl_cols   <- names(df)[str_ends(names(df), "_pcntl")]
score_cols   <- c("climate_score", "exposure_score", "effects_score",
                  "final_risk_score", "final_risk_score_pcntl")

cat(glue::glue("Dataset loaded: {n_facilities} facilities, {ncol(df)} columns\n"))

# WGS84 version for leaflet (centroids for point plotting)
df_wgs <- df_sf %>%
  st_transform(4326) %>%
  mutate(
    lon          = st_coordinates(st_centroid(st_geometry(.)))[,1],
    lat          = st_coordinates(st_centroid(st_geometry(.)))[,2],
    full_address = glue::glue("{ADDRESS}, {CITY}, {STATE} {ZIP}")
  )

# Shared risk color palette (percentile 0-100)
pal_risk <- colorNumeric("YlOrRd", domain = c(0, 100), na.color = "#aaaaaa")
```

------------------------------------------------------------------------

# Data Inventory & Quality Flags {.tabset}

## Completeness

```{r completeness}
completeness <- df %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "column", values_to = "n_missing") %>%
  mutate(pct_missing = n_missing / n_facilities) %>%
  arrange(desc(n_missing))

completeness %>%
  filter(n_missing > 0) %>%
  mutate(pct_missing = percent(pct_missing, accuracy = 0.1)) %>%
  kbl(caption = "Columns with any missing values") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  column_spec(3, bold = TRUE) %>%
  scroll_box(width = "100%", height = "250px")
```

## Range Checks

```{r range-checks}
range_spec <- tribble(
  ~column,             ~lo,  ~hi,   ~unit,
  "flood_risk",          0,  100,   "% flooded footprint",
  "wildfire_risk",       0,   NA,   "index (unbounded)",
  "lst_avg",            20,   60,   "°C (LST)",
  "percent_tree_cover",  0,  100,   "% canopy",
  "mean_ozone",         20,   90,   "ppb",
  "avg_pm25",            0,   50,   "µg/m³",
  "pesticide_mean_kg",   0,   NA,   "kg (unbounded)",
  "trafficProx",         0,   NA,   "IDW (unbounded)",
  "npl_prox",            0,   NA,   "inverse-distance score",
  "rmp_prox",            0,   NA,   "inverse-distance score",
  "haz_prox",            0,   NA,   "inverse-distance score"
)

range_results <- range_spec %>%
  rowwise() %>%
  mutate(
    n_valid  = sum(!is.na(df[[column]])),
    obs_min  = if (n_valid > 0) min(df[[column]], na.rm = TRUE) else NA_real_,
    obs_max  = if (n_valid > 0) max(df[[column]], na.rm = TRUE) else NA_real_
  ) %>%
  ungroup()

range_results %>%
  select(column, unit, n_valid, obs_min, obs_max) %>%
  mutate(across(c(obs_min, obs_max), ~ round(., 3))) %>%
  kbl(caption = "Range check for raw indicators") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  scroll_box(width = "100%", height = "250px")
```

## Facility Metadata

```{r metadata-type}
df %>%
  count(facility_type, name = "n_facilities") %>%
  arrange(desc(n_facilities)) %>%
  mutate(pct = percent(n_facilities / sum(n_facilities), accuracy = 0.1)) %>%
  kbl(caption = "Facilities by type (STATE vs FEDERAL)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  scroll_box(width = "100%", height = "250px")
```

```{r metadata-securelvl}
df %>%
  count(secure_level, name = "n_facilities") %>%
  arrange(desc(n_facilities)) %>%
  mutate(pct = percent(n_facilities / sum(n_facilities), accuracy = 0.1)) %>%
  kbl(caption = "Facilities by security level") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  scroll_box(width = "100%", height = "250px")
```

```{r metadata-state}
df %>%
  count(STATE, name = "n_facilities") %>%
  arrange(desc(n_facilities)) %>%
  mutate(pct = percent(n_facilities / sum(n_facilities), accuracy = 0.1)) %>%
  kbl(caption = "Facilities by state") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  scroll_box(width = "100%", height = "250px")
```

------------------------------------------------------------------------

# Raw Indicator Exploration

Note that for some indicators, values are designed to be interpreted as relative units only vs. holding actual metric value. You can view a breakdown of all data sources and indicator calculation methods here: <https://github.com/GeospatialCentroid/NASA-prison-EJ/blob/main/docs/indicator_metadata.md>

## Climate Indicators {.tabset}

### Flood Risk {.tabset}

#### Plot & Table

```{r flood}
p1 <- df %>%
  filter(!is.na(flood_risk)) %>%
  ggplot(aes(flood_risk)) +
  geom_histogram(bins = 40, fill = "#2c7fb8", color = "white", linewidth = 0.2) +
  geom_vline(xintercept = median(df$flood_risk, na.rm = TRUE),
             linetype = "dashed", color = "firebrick") +
  labs(title = "Flood Risk (% flood zone coverage)",
       subtitle = glue::glue("n = {sum(!is.na(df$flood_risk))} | {sum(is.na(df$flood_risk))} missing"),
       x = "% facility footprint + 1km buffer in flood zone", y = "count") +
  theme_minimal()

p2 <- df %>%
  ggplot(aes(x = "flood_risk", y = flood_risk)) +
  geom_boxplot(fill = "#2c7fb8", alpha = 0.5, outlier.colour = "firebrick") +
  labs(title = "Boxplot", x = NULL, y = NULL) +
  theme_minimal()

p1 + p2 + plot_layout(widths = c(3, 1))
```

```{r flood-flags}
flood_q90 <- quantile(df$flood_risk, 0.90, na.rm = TRUE)
df %>%
  filter(flood_risk >= flood_q90) %>%
  select(NAME, full_address, facility_type, secure_level, POPULATION, flood_risk) %>%
  arrange(desc(flood_risk)) %>%
  mutate(across(where(is.double), ~ round(., 2))) %>%
  kbl(caption = "Facilities above 90th Percentile — Flood Risk") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  scroll_box(width = "100%", height = "250px")
```

#### Map

**Facilities above 90th percentile flood risk**

```{r flood-map, out.width="100%", fig.height=4}
flood_flagged <- df_wgs %>% filter(flood_risk >= flood_q90)
pal_flood <- colorNumeric("Blues", domain = df_wgs$flood_risk, na.color = "#aaaaaa")

leaflet(flood_flagged, width = "100%", height = 400) %>%
  addProviderTiles(providers$CartoDB.Positron, group = "Carto Light") %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  setView(lng = -96, lat = 38, zoom = 4) %>%
  addLayersControl(
    baseGroups = c("Carto Light", "Satellite"),
    options = layersControlOptions(collapsed = TRUE)
  ) %>%
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    radius = 7,
    fillColor = ~pal_flood(flood_risk),
    fillOpacity = 0.85,
    color = "#1a478a", weight = 1,
    popup = ~glue::glue(
      "<b>{NAME}</b><br><i>{TYPE} | {SECURELVL}</i><br>{full_address}<hr>",
      "<b>Flood Risk:</b> {round(flood_risk, 1)}%<br>",
      "<b>Final Risk Score:</b> {round(final_risk_score, 2)}"
    ),
    label = ~NAME
  ) %>%
  addLegend("bottomright", pal = pal_flood, values = ~flood_risk,
            title = "Flood Risk (%)", opacity = 0.8)
```

------------------------------------------------------------------------

### Wildfire Risk {.tabset}

Note that these values are reflective of the [USFS Wildfire Hazard Potential Index](https://www.fs.usda.gov/rds/archive/catalog/RDS-2016-0034-3), which are meant to be relative values, not absolute.

#### Plot & Table

```{r wildfire}
p1 <- df %>%
  ggplot(aes(wildfire_risk)) +
  geom_histogram(bins = 40, fill = "#e34a33", color = "white", linewidth = 0.2) +
  scale_x_log10(labels = comma) +
  labs(title = "Wildfire Risk (log scale)",
       x = "wildfire risk index (log)", y = "count") +
  theme_minimal()

p2 <- df %>%
  ggplot(aes(x = "wildfire", y = wildfire_risk)) +
  geom_boxplot(fill = "#e34a33", alpha = 0.5, outlier.colour = "firebrick") +
  scale_y_log10(labels = comma) +
  labs(title = "Boxplot (log)", x = NULL, y = NULL) +
  theme_minimal()

p1 + p2 + plot_layout(widths = c(3, 1))
```

```{r wildfire-outliers}
wf_q90 <- quantile(df$wildfire_risk, 0.90, na.rm = TRUE)
df %>%
  filter(wildfire_risk > wf_q90) %>%
  select(NAME, full_address, facility_type, secure_level, wildfire_risk) %>%
  arrange(desc(wildfire_risk)) %>%
  mutate(across(where(is.double), ~ round(., 2))) %>%
  kbl(caption = glue::glue("Facilities above 90th percentile (>{round(wf_q90,1)})")) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  scroll_box(width = "100%", height = "250px")
```

#### Map

**Facilities above 90th percentile wildfire risk**

```{r wildfire-map, out.width="100%", fig.height=4}
wf_flagged <- df_wgs %>% filter(wildfire_risk > wf_q90)
pal_wf <- colorNumeric("Oranges", domain = log10(df_wgs$wildfire_risk + 1), na.color = "#aaaaaa")

leaflet(wf_flagged, width = "100%", height = 400) %>%
  addProviderTiles(providers$CartoDB.Positron, group = "Carto Light") %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  setView(lng = -96, lat = 38, zoom = 4) %>%
  addLayersControl(
    baseGroups = c("Carto Light", "Satellite"),
    options = layersControlOptions(collapsed = TRUE)
  ) %>%
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    radius = 7,
    fillColor = ~pal_wf(log10(wildfire_risk + 1)),
    fillOpacity = 0.85,
    color = "#b30000", weight = 1,
    popup = ~glue::glue(
      "<b>{NAME}</b><br><i>{TYPE} | {SECURELVL}</i><br>{full_address}<hr>",
      "<b>Wildfire Risk Index:</b> {round(wildfire_risk, 1)}<br>",
      "<b>Final Risk Score:</b> {round(final_risk_score, 2)}"
    ),
    label = ~NAME
  ) %>%
  addLegend("bottomright", pal = pal_wf, values = ~log10(wildfire_risk + 1),
            title = "Wildfire Risk (log₁₀)", opacity = 0.8)
```

------------------------------------------------------------------------

### Land Surface Temperature {.tabset}

#### Plot & Table

```{r lst}
p1 <- df %>%
  filter(!is.na(lst_avg)) %>%
  ggplot(aes(lst_avg)) +
  geom_histogram(bins = 40, fill = "#fd8d3c", color = "white", linewidth = 0.2) +
  geom_vline(xintercept = median(df$lst_avg, na.rm = TRUE),
             linetype = "dashed", color = "firebrick") +
  labs(title = "Mean LST (summer Landsat, °C)",
       subtitle = glue::glue("n = {sum(!is.na(df$lst_avg))} | {sum(is.na(df$lst_avg))} missing"),
       x = "LST °C", y = "count") +
  theme_minimal()

p1
```

```{r lst-flags}
lst_q90 <- quantile(df$lst_avg, 0.90, na.rm = TRUE)
df %>%
  filter(lst_avg > lst_q90) %>%
  select(NAME, full_address, facility_type, secure_level, lst_avg) %>%
  arrange(desc(lst_avg)) %>%
  mutate(across(where(is.double), ~ round(., 2))) %>%
  kbl(caption = glue::glue("Facilities above 90th percentile LST (>{round(lst_q90,1)} °C)")) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  scroll_box(width = "100%", height = "250px")
```

#### Map

**Facilities above 90th percentile land surface temperature**

```{r lst-map, out.width="100%", fig.height=4}
lst_flagged <- df_wgs %>% filter(lst_avg > lst_q90)
pal_lst <- colorNumeric("YlOrRd", domain = df_wgs$lst_avg, na.color = "#aaaaaa")

leaflet(lst_flagged, width = "100%", height = 400) %>%
  addProviderTiles(providers$CartoDB.Positron, group = "Carto Light") %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  setView(lng = -96, lat = 38, zoom = 4) %>%
  addLayersControl(
    baseGroups = c("Carto Light", "Satellite"),
    options = layersControlOptions(collapsed = TRUE)
  ) %>%
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    radius = 7,
    fillColor = ~pal_lst(lst_avg),
    fillOpacity = 0.85,
    color = "#7f0000", weight = 1,
    popup = ~glue::glue(
      "<b>{NAME}</b><br><i>{TYPE} | {SECURELVL}</i><br>{full_address}<hr>",
      "<b>Mean LST:</b> {round(lst_avg, 1)} °C<br>",
      "<b>Final Risk Score:</b> {round(final_risk_score, 2)}"
    ),
    label = ~NAME
  ) %>%
  addLegend("bottomright", pal = pal_lst, values = ~lst_avg,
            title = "Mean LST (°C)", opacity = 0.8)
```

------------------------------------------------------------------------

### Tree Canopy Cover {.tabset}

#### Plot & Table

```{r tree}
p1 <- df %>%
  filter(!is.na(percent_tree_cover)) %>%
  ggplot(aes(percent_tree_cover)) +
  geom_histogram(bins = 40, fill = "#41ab5d", color = "white", linewidth = 0.2) +
  labs(title = "Tree Canopy Cover (%)",
       x = "% canopy (1km buffer, NLCD 2023)", y = "count") +
  theme_minimal()

p1
```

```{r tree-flags}
df %>%
  filter(percent_tree_cover == 0) %>%
  select(NAME, full_address, facility_type, secure_level, percent_tree_cover) %>%
  mutate(across(where(is.double), ~ round(., 2))) %>%
  kbl(caption = "Facilities with 0% canopy") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  scroll_box(width = "100%", height = "250px")
```

#### Map

**Facilities with 0% Canopy Cover**

```{r tree-map, out.width="100%", fig.height=4}
leaflet(df_wgs %>% filter(percent_tree_cover == 0), width = "100%", height = 400) %>%
  addProviderTiles(providers$CartoDB.Positron, group = "Carto Light") %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  setView(lng = -96, lat = 38, zoom = 4) %>%
  addLayersControl(
    baseGroups = c("Carto Light", "Satellite"),
    options = layersControlOptions(collapsed = TRUE)
  ) %>%
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    radius = 7,
    fillColor = "#006d2c",
    fillOpacity = 0.85,
    color = "#006d2c", weight = 1,
    popup = ~glue::glue(
      "<b>{NAME}</b><br><i>{TYPE} | {SECURELVL}</i><br>{full_address}<hr>",
      "<b>Tree Canopy:</b> {round(percent_tree_cover, 1)}%<br>",
      "<b>Final Risk Score:</b> {round(final_risk_score, 2)}"
    ),
    label = ~NAME
  )
```

------------------------------------------------------------------------

## Exposure Indicators {.tabset}

### Ozone {.tabset}

#### Plot & Table

```{r ozone}
p1 <- df %>%
  ggplot(aes(mean_ozone)) +
  geom_histogram(bins = 40, fill = "#6a51a3", color = "white", linewidth = 0.2) +
  geom_vline(xintercept = 70, linetype = "dashed", color = "firebrick",
             linewidth = 0.8) +
  annotate("text", x = 71, y = Inf, label = "NAAQS 70 ppb",
           hjust = 0, vjust = 2, size = 3, color = "firebrick") +
  labs(title = "Mean 8-hr Max Ozone (ppb) — EPA FAQSD",
       x = "ozone (ppb)", y = "count") +
  theme_minimal()

p1
```

```{r ozone-flags}
n_exceed <- sum(df$mean_ozone >= 70, na.rm = TRUE)
cat(glue::glue("{n_exceed} facilities at or above NAAQS 70 ppb threshold\n\n"))

df %>%
  filter(mean_ozone >= 70) %>%
  select(NAME, full_address, facility_type, secure_level, mean_ozone) %>%
  arrange(desc(mean_ozone)) %>%
  mutate(across(where(is.double), ~ round(., 2))) %>%
  kbl(caption = "Facilities at or above NAAQS 70 ppb") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  scroll_box(width = "100%", height = "250px")
```

#### Map

**Facilities at or above NAAQS ozone standard (70 ppb)**

```{r ozone-map, out.width="100%", fig.height=4}
ozone_flagged <- df_wgs %>% filter(mean_ozone >= 70)
pal_ozone <- colorNumeric("Purples", domain = df_wgs$mean_ozone, na.color = "#aaaaaa")

leaflet(ozone_flagged, width = "100%", height = 400) %>%
  addProviderTiles(providers$CartoDB.Positron, group = "Carto Light") %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  setView(lng = -96, lat = 38, zoom = 4) %>%
  addLayersControl(
    baseGroups = c("Carto Light", "Satellite"),
    options = layersControlOptions(collapsed = TRUE)
  ) %>%
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    radius = 7,
    fillColor = ~pal_ozone(mean_ozone),
    fillOpacity = 0.85,
    color = "#3f007d", weight = 1,
    popup = ~glue::glue(
      "<b>{NAME}</b><br><i>{TYPE} | {SECURELVL}</i><br>{full_address}<hr>",
      "<b>Mean Ozone:</b> {round(mean_ozone, 1)} ppb<br>",
      "<b>NAAQS Standard:</b> 70 ppb<br>",
      "<b>Final Risk Score:</b> {round(final_risk_score, 2)}"
    ),
    label = ~NAME
  ) %>%
  addLegend("bottomright", pal = pal_ozone, values = ~mean_ozone,
            title = "Mean Ozone (ppb)", opacity = 0.8)
```

------------------------------------------------------------------------

### PM2.5 {.tabset}

#### Plot & Table

```{r pm25}
p1 <- df %>%
  ggplot(aes(avg_pm25)) +
  geom_histogram(bins = 40, fill = "#9ecae1", color = "white", linewidth = 0.2) +
  geom_vline(xintercept = 12, linetype = "dashed", color = "firebrick") +
  annotate("text", x = 12.5, y = Inf, label = "NAAQS 12 µg/m³",
           hjust = 0, vjust = 2, size = 3, color = "firebrick") +
  labs(title = "Mean PM2.5 (µg/m³)", x = "PM2.5 µg/m³", y = "count") +
  theme_minimal()

p1
```

```{r pm25-flags}
n_pm_exceed <- sum(df$avg_pm25 >= 12, na.rm = TRUE)
cat(glue::glue("{n_pm_exceed} facilities at or above annual NAAQS of 12 µg/m³\n\n"))

df %>%
  filter(avg_pm25 >= 12) %>%
  select(NAME, full_address, facility_type, secure_level, avg_pm25) %>%
  arrange(desc(avg_pm25)) %>%
  mutate(across(where(is.double), ~ round(., 2))) %>%
  kbl(caption = "Facilities at or above NAAQS 12 µg/m³") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  scroll_box(width = "100%", height = "250px")
```

#### Map

**Facilities at or above NAAQS annual PM2.5 standard (12 µg/m³)**

```{r pm25-map, out.width="100%", fig.height=4}
pm_flagged <- df_wgs %>% filter(avg_pm25 >= 12)
pal_pm <- colorNumeric("Blues", domain = df_wgs$avg_pm25, na.color = "#aaaaaa")

leaflet(pm_flagged, width = "100%", height = 400) %>%
  addProviderTiles(providers$CartoDB.Positron, group = "Carto Light") %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  setView(lng = -96, lat = 38, zoom = 4) %>%
  addLayersControl(
    baseGroups = c("Carto Light", "Satellite"),
    options = layersControlOptions(collapsed = TRUE)
  ) %>%
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    radius = 7,
    fillColor = ~pal_pm(avg_pm25),
    fillOpacity = 0.85,
    color = "#08306b", weight = 1,
    popup = ~glue::glue(
      "<b>{NAME}</b><br><i>{TYPE} | {SECURELVL}</i><br>{full_address}<hr>",
      "<b>Mean PM2.5:</b> {round(avg_pm25, 2)} µg/m³<br>",
      "<b>NAAQS Standard:</b> 12 µg/m³<br>",
      "<b>Final Risk Score:</b> {round(final_risk_score, 2)}"
    ),
    label = ~NAME
  ) %>%
  addLegend("bottomright", pal = pal_pm, values = ~avg_pm25,
            title = "PM2.5 (µg/m³)", opacity = 0.8)
```

------------------------------------------------------------------------

### Pesticide Load {.tabset}

#### Plot & Table

```{r pesticide}
p1 <- df %>%
  filter(!is.na(pesticide_mean_kg)) %>%
  ggplot(aes(pesticide_mean_kg)) +
  geom_histogram(bins = 40, fill = "#a1d99b", color = "white", linewidth = 0.2) +
  scale_x_log10(labels = comma) +
  labs(title = "Mean Pesticide Load (kg, PEST-CHEMGRIDSv2, log scale)",
       subtitle = glue::glue("n = {sum(!is.na(df$pesticide_mean_kg))} | ",
                             "{sum(df$pesticide_mean_kg == 0, na.rm=TRUE)} zeros"),
       x = "pesticide mean kg (log₁₀)", y = "count") +
  theme_minimal()

p2 <- df %>%
  filter(!is.na(pesticide_mean_kg)) %>%
  ggplot(aes("pesticide", pesticide_mean_kg)) +
  geom_boxplot(fill = "#a1d99b", alpha = 0.5, outlier.colour = "firebrick") +
  scale_y_log10(labels = comma) +
  labs(title = "Boxplot (log)", x = NULL, y = NULL) +
  theme_minimal()

p1 + p2 + plot_layout(widths = c(3, 1))
```

```{r pesticide-flags}
pest_q90 <- quantile(df$pesticide_mean_kg, 0.90, na.rm = TRUE)
df %>%
  filter(pesticide_mean_kg >= pest_q90) %>%
  select(NAME, full_address, facility_type, secure_level, pesticide_mean_kg) %>%
  arrange(desc(pesticide_mean_kg)) %>%
  mutate(across(where(is.double), ~ round(., 1))) %>%
  kbl(caption = glue::glue("Facilities above 90th percentile pesticide load (>{round(pest_q90,1)} kg)")) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  scroll_box(width = "100%", height = "250px")
```

#### Map

**Facilities above 90th percentile pesticide load**

```{r pesticide-map, out.width="100%", fig.height=4}
pest_flagged <- df_wgs %>% filter(pesticide_mean_kg >= pest_q90)
pal_pest <- colorNumeric("Greens", domain = log10(df_wgs$pesticide_mean_kg + 1), na.color = "#aaaaaa")

leaflet(pest_flagged, width = "100%", height = 400) %>%
  addProviderTiles(providers$CartoDB.Positron, group = "Carto Light") %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  setView(lng = -96, lat = 38, zoom = 4) %>%
  addLayersControl(
    baseGroups = c("Carto Light", "Satellite"),
    options = layersControlOptions(collapsed = TRUE)
  ) %>%
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    radius = 7,
    fillColor = ~pal_pest(log10(pesticide_mean_kg + 1)),
    fillOpacity = 0.85,
    color = "#00441b", weight = 1,
    popup = ~glue::glue(
      "<b>{NAME}</b><br><i>{TYPE} | {SECURELVL}</i><br>{full_address}<hr>",
      "<b>Pesticide Load:</b> {round(pesticide_mean_kg, 1)} kg<br>",
      "<b>Final Risk Score:</b> {round(final_risk_score, 2)}"
    ),
    label = ~NAME
  ) %>%
  addLegend("bottomright", pal = pal_pest, values = ~log10(pesticide_mean_kg + 1),
            title = "Pesticide (log₁₀ kg)", opacity = 0.8)
```

------------------------------------------------------------------------

### Traffic Proximity {.tabset}

#### Plot & Table

```{r traffic}
p1 <- df %>%
  ggplot(aes(trafficProx)) +
  geom_histogram(bins = 40, fill = "#636363", color = "white", linewidth = 0.2) +
  scale_x_log10(labels = comma) +
  labs(title = "Traffic Proximity Score (IDW, 500m buffer, log scale)",
       x = "traffic proximity (log₁₀)", y = "count") +
  theme_minimal()

p2 <- df %>%
  ggplot(aes("traffic", trafficProx)) +
  geom_boxplot(fill = "#636363", alpha = 0.5, outlier.colour = "firebrick") +
  scale_y_log10(labels = comma) +
  labs(title = "Boxplot (log)", x = NULL, y = NULL) +
  theme_minimal()

p1 + p2 + plot_layout(widths = c(3, 1))
```

```{r traffic-flags}
traffic_q90 <- quantile(df$trafficProx, 0.90, na.rm = TRUE)
df %>%
  filter(trafficProx >= traffic_q90) %>%
  select(NAME, full_address, facility_type, secure_level, trafficProx) %>%
  arrange(desc(trafficProx)) %>%
  mutate(across(where(is.double), ~ round(., 2))) %>%
  kbl(caption = "Facilities above 90th percentile — Traffic Proximity") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  scroll_box(width = "100%", height = "250px")
```

#### Map

**Facilities above 90th percentile traffic proximity score**

```{r traffic-map, out.width="100%", fig.height=4}
traffic_flagged <- df_wgs %>% filter(trafficProx >= traffic_q90)
pal_traffic <- colorNumeric("Greys", domain = log10(df_wgs$trafficProx + 1), na.color = "#aaaaaa")

leaflet(traffic_flagged, width = "100%", height = 400) %>%
  addProviderTiles(providers$CartoDB.Positron, group = "Carto Light") %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  setView(lng = -96, lat = 38, zoom = 4) %>%
  addLayersControl(
    baseGroups = c("Carto Light", "Satellite"),
    options = layersControlOptions(collapsed = TRUE)
  ) %>%
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    radius = 7,
    fillColor = ~pal_traffic(log10(trafficProx + 1)),
    fillOpacity = 0.85,
    color = "#252525", weight = 1,
    popup = ~glue::glue(
      "<b>{NAME}</b><br><i>{TYPE} | {SECURELVL}</i><br>{full_address}<hr>",
      "<b>Traffic Proximity Score:</b> {round(trafficProx, 1)}<br>",
      "<b>Final Risk Score:</b> {round(final_risk_score, 2)}"
    ),
    label = ~NAME
  ) %>%
  addLegend("bottomright", pal = pal_traffic, values = ~log10(trafficProx + 1),
            title = "Traffic Proximity (log₁₀)", opacity = 0.8)
```

------------------------------------------------------------------------

## Effects Indicators {.tabset}

### NPL Proximity {.tabset}

#### Plot & Table

```{r npl}
p1 <- df %>%
  ggplot(aes(npl_prox)) +
  geom_histogram(bins = 40, fill = "#2ca25f", color = "white", linewidth = 0.2) +
  scale_x_log10(labels = comma) +
  labs(title = "NPL (Superfund) Site Proximity Score (log scale)",
       x = "NPL proximity score (log₁₀)", y = "count") +
  theme_minimal()

p2 <- df %>%
  ggplot(aes("npl", npl_prox)) +
  geom_boxplot(fill = "#2ca25f", alpha = 0.5, outlier.colour = "firebrick") +
  scale_y_log10(labels = comma) +
  labs(title = "Boxplot (log)", x = NULL, y = NULL) +
  theme_minimal()

p1 + p2 + plot_layout(widths = c(3, 1))
```

```{r npl-flags}
npl_q90 <- quantile(df$npl_prox, 0.90, na.rm = TRUE)
df %>%
  filter(npl_prox >= npl_q90) %>%
  select(NAME, full_address, facility_type, secure_level, npl_prox) %>%
  arrange(desc(npl_prox)) %>%
  mutate(across(where(is.double), ~ round(., 3))) %>%
  kbl(caption = "Facilities above 90th percentile — NPL Proximity") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  scroll_box(width = "100%", height = "250px")
```

#### Map

**Facilities above 90th percentile NPL (Superfund) site proximity**

```{r npl-map, out.width="100%", fig.height=4}
npl_flagged <- df_wgs %>% filter(npl_prox >= npl_q90)
pal_npl <- colorNumeric("YlGn", domain = log10(df_wgs$npl_prox + 1), na.color = "#aaaaaa")

leaflet(npl_flagged, width = "100%", height = 400) %>%
  addProviderTiles(providers$CartoDB.Positron, group = "Carto Light") %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  setView(lng = -96, lat = 38, zoom = 4) %>%
  addLayersControl(
    baseGroups = c("Carto Light", "Satellite"),
    options = layersControlOptions(collapsed = TRUE)
  ) %>%
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    radius = 7,
    fillColor = ~pal_npl(log10(npl_prox + 1)),
    fillOpacity = 0.85,
    color = "#006d2c", weight = 1,
    popup = ~glue::glue(
      "<b>{NAME}</b><br><i>{TYPE} | {SECURELVL}</i><br>{full_address}<hr>",
      "<b>NPL Proximity Score:</b> {round(npl_prox, 3)}<br>",
      "<b>Final Risk Score:</b> {round(final_risk_score, 2)}"
    ),
    label = ~NAME
  ) %>%
  addLegend("bottomright", pal = pal_npl, values = ~log10(npl_prox + 1),
            title = "NPL Proximity (log₁₀)", opacity = 0.8)
```

------------------------------------------------------------------------

### RMP Proximity {.tabset}

#### Plot & Table

```{r rmp}
p1 <- df %>%
  ggplot(aes(rmp_prox)) +
  geom_histogram(bins = 40, fill = "#f768a1", color = "white", linewidth = 0.2) +
  scale_x_log10(labels = comma) +
  labs(title = "RMP Site Proximity Score (log scale)",
       x = "RMP proximity score (log₁₀)", y = "count") +
  theme_minimal()

p2 <- df %>%
  ggplot(aes("rmp", rmp_prox)) +
  geom_boxplot(fill = "#f768a1", alpha = 0.5, outlier.colour = "firebrick") +
  scale_y_log10(labels = comma) +
  labs(title = "Boxplot (log)", x = NULL, y = NULL) +
  theme_minimal()

p1 + p2 + plot_layout(widths = c(3, 1))
```

```{r rmp-flags}
rmp_q90 <- quantile(df$rmp_prox, 0.90, na.rm = TRUE)
df %>%
  filter(rmp_prox >= rmp_q90) %>%
  select(NAME, full_address, facility_type, secure_level, rmp_prox) %>%
  arrange(desc(rmp_prox)) %>%
  mutate(across(where(is.double), ~ round(., 3))) %>%
  kbl(caption = "Facilities above 90th percentile — RMP Proximity") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  scroll_box(width = "100%", height = "250px")
```

#### Map

**Facilities above 90th percentile RMP site proximity**

```{r rmp-map, out.width="100%", fig.height=4}
rmp_flagged <- df_wgs %>% filter(rmp_prox >= rmp_q90)
pal_rmp <- colorNumeric("RdPu", domain = log10(df_wgs$rmp_prox + 1), na.color = "#aaaaaa")

leaflet(rmp_flagged, width = "100%", height = 400) %>%
  addProviderTiles(providers$CartoDB.Positron, group = "Carto Light") %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  setView(lng = -96, lat = 38, zoom = 4) %>%
  addLayersControl(
    baseGroups = c("Carto Light", "Satellite"),
    options = layersControlOptions(collapsed = TRUE)
  ) %>%
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    radius = 7,
    fillColor = ~pal_rmp(log10(rmp_prox + 1)),
    fillOpacity = 0.85,
    color = "#7a0177", weight = 1,
    popup = ~glue::glue(
      "<b>{NAME}</b><br><i>{TYPE} | {SECURELVL}</i><br>{full_address}<hr>",
      "<b>RMP Proximity Score:</b> {round(rmp_prox, 3)}<br>",
      "<b>Final Risk Score:</b> {round(final_risk_score, 2)}"
    ),
    label = ~NAME
  ) %>%
  addLegend("bottomright", pal = pal_rmp, values = ~log10(rmp_prox + 1),
            title = "RMP Proximity (log₁₀)", opacity = 0.8)
```

------------------------------------------------------------------------

### Hazardous Waste Proximity {.tabset}

#### Plot & Table

```{r haz}
p1 <- df %>%
  ggplot(aes(haz_prox)) +
  geom_histogram(bins = 40, fill = "#ae017e", color = "white", linewidth = 0.2) +
  scale_x_log10(labels = comma) +
  labs(title = "Hazardous Waste Site Proximity Score (log scale)",
       x = "haz proximity score (log₁₀)", y = "count") +
  theme_minimal()

p2 <- df %>%
  ggplot(aes("haz", haz_prox)) +
  geom_boxplot(fill = "#ae017e", alpha = 0.5, outlier.colour = "firebrick") +
  scale_y_log10(labels = comma) +
  labs(title = "Boxplot (log)", x = NULL, y = NULL) +
  theme_minimal()

p1 + p2 + plot_layout(widths = c(3, 1))
```

```{r haz-flags}
haz_q90 <- quantile(df$haz_prox, 0.90, na.rm = TRUE)
df %>%
  filter(haz_prox >= haz_q90) %>%
  select(NAME, full_address, facility_type, secure_level, haz_prox) %>%
  arrange(desc(haz_prox)) %>%
  mutate(across(where(is.double), ~ round(., 3))) %>%
  kbl(caption = "Facilities above 90th percentile — Hazardous Waste Proximity") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  scroll_box(width = "100%", height = "250px")
```

#### Map

**Facilities above 90th percentile hazardous waste site proximity**

```{r haz-map, out.width="100%", fig.height=4}
haz_flagged <- df_wgs %>% filter(haz_prox >= haz_q90)
pal_haz <- colorNumeric("Reds", domain = log10(df_wgs$haz_prox + 1), na.color = "#aaaaaa")

leaflet(haz_flagged, width = "100%", height = 400) %>%
  addProviderTiles(providers$CartoDB.Positron, group = "Carto Light") %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  setView(lng = -96, lat = 38, zoom = 4) %>%
  addLayersControl(
    baseGroups = c("Carto Light", "Satellite"),
    options = layersControlOptions(collapsed = TRUE)
  ) %>%
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    radius = 7,
    fillColor = ~pal_haz(log10(haz_prox + 1)),
    fillOpacity = 0.85,
    color = "#67000d", weight = 1,
    popup = ~glue::glue(
      "<b>{NAME}</b><br><i>{TYPE} | {SECURELVL}</i><br>{full_address}<hr>",
      "<b>Hazardous Waste Proximity Score:</b> {round(haz_prox, 3)}<br>",
      "<b>Final Risk Score:</b> {round(final_risk_score, 2)}"
    ),
    label = ~NAME
  ) %>%
  addLegend("bottomright", pal = pal_haz, values = ~log10(haz_prox + 1),
            title = "Haz Waste Proximity (log₁₀)", opacity = 0.8)
```

------------------------------------------------------------------------

# Component & Index-Level Summaries {.tabset}

## Score Correlations

```{r score-corr}
component_scores <- c("climate_score", "exposure_score", "effects_score")

cor_mat <- df %>%
  select(all_of(component_scores), final_risk_score) %>%
  cor(use = "pairwise.complete.obs", method = "spearman")

cor_long <- as.data.frame(as.table(cor_mat)) %>%
  filter(as.character(Var1) < as.character(Var2)) %>%
  rename(score_a = Var1, score_b = Var2, spearman_r = Freq)

cor_long %>%
  mutate(spearman_r = round(spearman_r, 3)) %>%
  kbl(caption = "Spearman rank correlations between component scores and final index") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  scroll_box(width = "100%", height = "250px")
```

## Scores by Facility Type

```{r scores-by-type}
df %>%
  select(facility_type, all_of(component_scores), final_risk_score) %>%
  pivot_longer(-facility_type, names_to = "score", values_to = "value") %>%
  mutate(score = factor(score,
                        levels = c("climate_score", "exposure_score",
                                   "effects_score", "final_risk_score"))) %>%
  ggplot(aes(x = facility_type, y = value, fill = facility_type)) +
  geom_boxplot(alpha = 0.7, outlier.size = 1) +
  facet_wrap(~ score, ncol = 2, scales = "free_y") +
  labs(title = "Component and Final Scores by Facility Type (STATE vs FEDERAL)",
       x = NULL, y = "score") +
  theme_minimal() +
  theme(legend.position = "none")
```

## Scores by Security Level

```{r scores-by-securelvl}
df %>%
  filter(SECURELVL != "NOT AVAILABLE") %>%
  select(secure_level, all_of(component_scores), final_risk_score) %>%
  pivot_longer(-secure_level, names_to = "score", values_to = "value") %>%
  mutate(score = factor(score,
                        levels = c("climate_score", "exposure_score",
                                   "effects_score", "final_risk_score"))) %>%
  ggplot(aes(x = reorder(secure_level, value, median, na.rm = TRUE),
             y = value, fill = secure_level)) +
  geom_boxplot(alpha = 0.7, outlier.size = 1) +
  coord_flip() +
  facet_wrap(~ score, ncol = 2, scales = "free_x") +
  labs(title = "Component and Final Scores by Security Level",
       x = NULL, y = "score") +
  theme_minimal() +
  theme(legend.position = "none")
```

## Top/Bottom Facilities (Final Index)

```{r top-facilities}
top_n_show <- 20

df %>%
  arrange(desc(final_risk_score_pcntl)) %>%
  slice_head(n = top_n_show) %>%
  select(NAME, facility_type, full_address, STATE, SECURELVL,
         POPULATION, climate_score, exposure_score, effects_score,
         final_risk_score, final_risk_score_pcntl) %>%
  mutate(across(where(is.double), ~ round(., 2))) %>%
  kbl(caption = glue::glue("Top {top_n_show} highest-risk facilities (final index percentile)")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                font_size = 12, full_width = FALSE) %>%
  row_spec(1:5, background = "#fff3cd") %>%
  scroll_box(width = "100%", height = "250px")
```

```{r bottom-facilities}
df %>%
  arrange(final_risk_score_pcntl) %>%
  slice_head(n = top_n_show) %>%
  select(NAME, facility_type, full_address, STATE, SECURELVL,
         POPULATION, climate_score, exposure_score, effects_score,
         final_risk_score, final_risk_score_pcntl) %>%
  mutate(across(where(is.double), ~ round(., 2))) %>%
  kbl(caption = glue::glue("Bottom {top_n_show} lowest-risk facilities")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                font_size = 12, full_width = FALSE) %>%
  scroll_box(width = "100%", height = "250px")
```

## Indicator Heatmap (Top 40 Facilities)

```{r heatmap, fig.height=11, fig.width=10}
pcntl_raw <- df %>%
  select(NAME, ends_with("_pcntl"), -final_risk_score_pcntl) %>%
  arrange(desc(df$final_risk_score_pcntl)) %>%
  slice_head(n = 40) %>%
  pivot_longer(-NAME, names_to = "indicator", values_to = "percentile") %>%
  mutate(
    indicator = str_remove(indicator, "_pcntl"),
    indicator = str_replace_all(indicator, "_", " ")
  )

ggplot(pcntl_raw, aes(x = indicator,
                      y = reorder(NAME, percentile, mean, na.rm = TRUE),
                      fill = percentile)) +
  geom_tile(color = "white", linewidth = 0.3) +
  scale_fill_gradient2(low = "#2166ac", mid = "#f7f7f7", high = "#d6604d",
                       midpoint = 50, limits = c(0, 100),
                       name = "Percentile") +
  labs(title = "Indicator Percentile Heatmap — Top 40 Highest-Risk Facilities",
       subtitle = "Sorted by mean percentile across indicators; red = high exposure",
       x = NULL, y = NULL) +
  theme_minimal() +
  theme(
    axis.text.x  = element_text(angle = 40, hjust = 1, size = 9),
    axis.text.y  = element_text(size = 7),
    panel.grid   = element_blank()
  )
```

------------------------------------------------------------------------

# Spatial Patterns {.tabset}

## Final Risk Index — All Facilities

All `r nrow(df_wgs)` facilities mapped and colored by final risk score. Click any marker for full details.

```{r map-final-risk, out.width="100%", fig.height=4}
leaflet(df_wgs, width = "100%", height = 450) %>%
  addProviderTiles(providers$CartoDB.Positron, group = "Carto Light") %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  setView(lng = -96, lat = 38, zoom = 4) %>%
  addLayersControl(
    baseGroups = c("Carto Light", "Satellite"),
    options = layersControlOptions(collapsed = TRUE)
  ) %>%
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    radius = ~scales::rescale(final_risk_score, to = c(4, 12)),
    fillColor = ~pal_risk(final_risk_score_pcntl),
    fillOpacity = 0.85,
    color = "#444444", weight = 0.5,
    popup = ~glue::glue(
      "<b>{NAME}</b><br><i>{TYPE} | {SECURELVL}</i><br>{full_address}<hr>",
      "<b>Final Risk Percentile:</b> {round(final_risk_score_pcntl, 1)}<br>",
      "<b>Climate Score:</b> {round(climate_score, 2)}<br>",
      "<b>Exposure Score:</b> {round(exposure_score, 2)}<br>",
      "<b>Effects Score:</b> {round(effects_score, 2)}<br>",
      "<b>Final Risk Score:</b> {round(final_risk_score, 2)}<br>",
      "<b>Population:</b> {scales::comma(POPULATION)}"
    ),
    label = ~NAME
  ) %>%
  addLegend("bottomright", pal = pal_risk,
            values = ~final_risk_score_pcntl,
            title = "Final Risk Percentile", opacity = 0.85)
```

------------------------------------------------------------------------

## Top 25 Highest-Risk Facilities

```{r map-top25, out.width="100%", fig.height=4}
top25 <- df_wgs %>%
  arrange(desc(final_risk_score_pcntl)) %>%
  slice_head(n = 25) %>%
  mutate(rank = row_number())

pal_top25 <- colorNumeric("YlOrRd", domain = top25$final_risk_score_pcntl)

leaflet(top25, width = "100%", height = 450) %>%
  addProviderTiles(providers$CartoDB.Positron, group = "Carto Light") %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  setView(lng = -96, lat = 38, zoom = 4) %>%
  addLayersControl(
    baseGroups = c("Carto Light", "Satellite"),
    options = layersControlOptions(collapsed = TRUE)
  ) %>%
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    radius = 9,
    fillColor = ~pal_top25(final_risk_score_pcntl),
    fillOpacity = 0.9,
    color = "#222222", weight = 1,
    popup = ~glue::glue(
      "<b>Rank #{rank}: {NAME}</b><br><i>{TYPE} | {SECURELVL}</i><br>{full_address}<hr>",
      "<b>Final Risk Percentile:</b> {round(final_risk_score_pcntl, 1)}<br>",
      "<b>Climate Score:</b> {round(climate_score, 2)}<br>",
      "<b>Exposure Score:</b> {round(exposure_score, 2)}<br>",
      "<b>Effects Score:</b> {round(effects_score, 2)}<br>",
      "<b>Population:</b> {scales::comma(POPULATION)}<br>",
      "<b>Capacity:</b> {scales::comma(CAPACITY)}"
    ),
    label = ~glue::glue("#{rank} {NAME}"),
    labelOptions = labelOptions(permanent = FALSE)
  ) %>%
  addLegend("bottomright", pal = pal_top25,
            values = ~final_risk_score_pcntl,
            title = "Final Risk Percentile", opacity = 0.85)
```

------------------------------------------------------------------------

## Component Score Maps {.tabset}

### Climate

```{r map-climate, out.width="100%", fig.height=4}
pal_climate <- colorNumeric("YlOrRd", domain = df_wgs$climate_score, na.color = "#aaaaaa")

leaflet(df_wgs, width = "100%", height = 450) %>%
  addProviderTiles(providers$CartoDB.Positron, group = "Carto Light") %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  setView(lng = -96, lat = 38, zoom = 4) %>%
  addLayersControl(
    baseGroups = c("Carto Light", "Satellite"),
    options = layersControlOptions(collapsed = TRUE)
  ) %>%
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    radius = 6,
    fillColor = ~pal_climate(climate_score),
    fillOpacity = 0.8,
    color = "#555555", weight = 0.5,
    popup = ~glue::glue(
      "<b>{NAME}</b><br><i>{TYPE} | {SECURELVL}</i><br>{full_address}<hr>",
      "<b>Climate Score:</b> {round(climate_score, 2)}<br>",
      "<b>Flood Risk:</b> {round(flood_risk, 1)}%<br>",
      "<b>Wildfire Risk:</b> {round(wildfire_risk, 1)}<br>",
      "<b>LST:</b> {round(lst_avg, 1)} °C<br>",
      "<b>Tree Canopy:</b> {round(percent_tree_cover, 1)}%"
    ),
    label = ~NAME
  ) %>%
  addLegend("bottomright", pal = pal_climate,
            values = ~climate_score,
            title = "Climate Score", opacity = 0.85)
```

### Exposure

```{r map-exposure, out.width="100%", fig.height=4}
pal_exposure <- colorNumeric("YlOrRd", domain = df_wgs$exposure_score, na.color = "#aaaaaa")

leaflet(df_wgs, width = "100%", height = 450) %>%
  addProviderTiles(providers$CartoDB.Positron, group = "Carto Light") %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  setView(lng = -96, lat = 38, zoom = 4) %>%
  addLayersControl(
    baseGroups = c("Carto Light", "Satellite"),
    options = layersControlOptions(collapsed = TRUE)
  ) %>%
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    radius = 6,
    fillColor = ~pal_exposure(exposure_score),
    fillOpacity = 0.8,
    color = "#555555", weight = 0.5,
    popup = ~glue::glue(
      "<b>{NAME}</b><br><i>{TYPE} | {SECURELVL}</i><br>{full_address}<hr>",
      "<b>Exposure Score:</b> {round(exposure_score, 2)}<br>",
      "<b>Ozone:</b> {round(mean_ozone, 1)} ppb<br>",
      "<b>PM2.5:</b> {round(avg_pm25, 2)} µg/m³<br>",
      "<b>Pesticide Load:</b> {round(pesticide_mean_kg, 1)} kg<br>",
      "<b>Traffic Proximity:</b> {round(trafficProx, 1)}"
    ),
    label = ~NAME
  ) %>%
  addLegend("bottomright", pal = pal_exposure,
            values = ~exposure_score,
            title = "Exposure Score", opacity = 0.85)
```

### Effects

```{r map-effects, out.width="100%", fig.height=4}
pal_effects <- colorNumeric("YlOrRd", domain = df_wgs$effects_score, na.color = "#aaaaaa")

leaflet(df_wgs, width = "100%", height = 450) %>%
  addProviderTiles(providers$CartoDB.Positron, group = "Carto Light") %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  setView(lng = -96, lat = 38, zoom = 4) %>%
  addLayersControl(
    baseGroups = c("Carto Light", "Satellite"),
    options = layersControlOptions(collapsed = TRUE)
  ) %>%
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    radius = 6,
    fillColor = ~pal_effects(effects_score),
    fillOpacity = 0.8,
    color = "#555555", weight = 0.5,
    popup = ~glue::glue(
      "<b>{NAME}</b><br><i>{TYPE} | {SECURELVL}</i><br>{full_address}<hr>",
      "<b>Effects Score:</b> {round(effects_score, 2)}<br>",
      "<b>NPL Proximity:</b> {round(npl_prox, 3)}<br>",
      "<b>RMP Proximity:</b> {round(rmp_prox, 3)}<br>",
      "<b>Haz Waste Proximity:</b> {round(haz_prox, 3)}"
    ),
    label = ~NAME
  ) %>%
  addLegend("bottomright", pal = pal_effects,
            values = ~effects_score,
            title = "Effects Score", opacity = 0.85)
```

------------------------------------------------------------------------
